---
title: "W4 Course Project: Exercise manner predict"
author: "WangYong"
date: "2022/2/3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Executive Summary
In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of **6 participants**. They were asked to perform barbell lifts correctly and incorrectly in **5 different ways**. The goal of your project is to predict the manner in which they did the exercise. 

This is the "classe" variable in the training set. You may use any of the other variables to predict with any of the other variables to predict with. 

You should create a report describing 
- how you built your model, 
- how you used cross validation, 
- what you think the expected out of sample error is, and 
- why you made the choices you did. 
- predict 20 different test cases. 

# 1. Exploratory Data Analysis
```{r init, echo=TRUE,message=FALSE, warning=FALSE}
library(dplyr) # data manipulate
library(caret) # machine learning
library(ggplot2) # data visualization
library(data.table) # data loading
library(lubridate) # datetime manipulate
library(moments)  # caculate moments such skew, and etc.
library(mice)  # missing value imputation
library(rattle) # fancyplot for decision treee
```
## 1.1 Load Data  
load the pml-training.csv as pml-train, pml-testing.csv as pml-test. 
```{r load, echo=FALSE,message=FALSE, warning=FALSE,cache=TRUE}
pml_train <- fread('pml-training.csv')
pml_test <- fread('pml-training.csv')
dim(pml_train)

```
The original data has 19622 obs, and 160 variables. 

## 1.2 check the outcome by contingency table
```{r,eda,echo=FALSE}
table(pml_train$user_name,pml_train$classe)
```
It can be find there is 6 people and 5 types of exercise classes(outcome, need predict) 

# 2. PreProcess

## 2.1 drop missing value variable and keep 60 variable
```{r mice, echo=FALSE}
library(ggplot2)
drop_ratio=0.97
get_missing_rate<-function(x) { 100*sum(is.na(x))/length(x)}
var_na_scores <- sapply(pml_train, get_missing_rate) 
na_rate_df <- data.frame(table(round(var_na_scores,1)))
g<- ggplot(data=na_rate_df,aes(Var1, Freq))
g+geom_col(fill='blue' ) +
  #coord_polar("y", start=0)+
  coord_flip()+
  geom_text(aes(label=Freq), position=position_dodge(0.9), hjust=-0.5)+
  labs(  x='the rate(%) of missing',
         y='#the number of variables',
         title='Missing Value Variables status'
         )
# drop the 100% missing value variable, they 
var_dropped <-names(var_na_scores[var_na_scores>100*drop_ratio])
pre_train <- pml_train %>% select(-all_of(var_dropped))
pre_test <- pml_test %>% select(-all_of(var_dropped))
#print(var_na_full)
```

The loaded data in pml-training has 160 variable. However, 100 variabls are 
almost missing value, the missing value rate is more than 97%. All of the 100 variables
will be dropped. Only 60 of them will be kept.

## 2.2 transform datetime related variables
There are 3 variable related to date time, after comparing, we translate them into
day, weekday, hour, minutes.
```{r trans_time,echo =FALSE}
pre_train<- pre_train %>% 
  select(-raw_timestamp_part_1, -raw_timestamp_part_2)%>%
  mutate(dt=dmy_hm(cvtd_timestamp)) %>%
  mutate(day=day(dt),wday=wday(dt),hour=hour(dt),minute=minute(dt))%>%  
  select( -cvtd_timestamp,-dt) 
#%>%mutate(user_name=factor(user_name),new_window=factor(new_window))
pre_test<- pre_test %>% 
   select(-raw_timestamp_part_1, -raw_timestamp_part_2)%>%
  mutate(dt=dmy_hm(cvtd_timestamp)) %>%
  mutate(day=day(dt),wday=wday(dt),hour=hour(dt),minute=minute(dt))%>%
  select(  -cvtd_timestamp,-dt) 
#%>%  mutate(user_name=factor(user_name),new_window=factor(new_window))
```


## 2.3 check variable skewness
```{r logtrans, echo=FASLSE}
get_log_df<-function(df){
  skew_criteria <- 5
  char_vars <- df %>% select_if(is.character) %>% names()
  numeric_var <- df %>% select_if(is.numeric) %>% names()

  log_vars<- df %>% 
    summarize(across(all_of(numeric_var),skewness)) %>% t()%>%  
    data.frame(score=.) 
  log_vars
}
get_log_df(pre_train) %>% filter(abs(.)>2) %>% arrange(desc(abs(score)))
  
```
According to above figure, we can find there is some variable is highly skewed. 
For example, gyros_dumbbel_z(135), gyros_dumbbel_x(-126), and etc.
Later, we will use preProcess Boxcox to transform which can transform negative
value as well.

# 2.4 All in one preProcess 
In this stage, the train/test dataset will be preProcess all in one style:
- 1. nzv remove near zero values
- 2. BoxCox, transform skewed variable
- 3. center,
- 4. scale, 

```{r mice,echo=TRUE, message=FALSE, warning=FALSE}
pre_train <- pml_train %>% select(-all_of(var_na_full)) 
pre_test  <- pml_test  %>% select(-all_of(var_na_full)) 
ml_pre <- preProcess(pre_train, 
                     method= c("nzv","BoxCox","center", "scale")
                     )
pre_train <- predict(ml_pre, pre_train)
pre_test <- predict(ml_pre, pre_test)
print(ml_pre)
```
Now, we has finished the preProcess stage. the training data set has reduced from
160 variable to 60 variables(include the outcome variable 'classe').


# 3. Data Slice
According to the best practice, the train dataset is medium size, so we slice it 60% as training and 40% as validation.
```{r slide, echo=FALSE}
in_train <- createDataPartition(y=pre_train$classe,p=0.6,list=FALSE)
train_df <-pre_train[in_train,]
validate_df <- pre_train[-in_train,]
```
Now, we has slide the raw pml_train into two part:   
- train_df, which as `r dim(train_df)[1]` observation.
- validation_df, which as `r dim(validation_df)[1]` observation.


# 4. Train
## 4.1 base mode
In traing stage, we will choice 'rpart' as the train method. our goal is 
classification which predict 1 of 5 movement classes based on data. 
```{r train,cache=TRUE}
set.seed(42)
fit_base <- train(data=train_df,
                classe~., 
                method='rpart',
                )
fit_base$results[2][1,]
```
In the base mode, we got 74% accuracy mode.
```{r,fancyplot, echo=FALSE}
fancyRpartPlot(fit_base$finalModel)
```
# Validatioin with ConfusionMatrix
```{r confusionMatrix}
set.seed(42)

pred <- predict(fit_base, validate_df)

confusionMatrix(factor(pred), factor(validate_df$classe))
```

# think the expected out of sample error

#  why you made the choices you did

# Predict

## Including Plots
